\newpage
\section{Configuration}

The Generalized Orthogonal Vectors Generator and Visualizer package provides a flexible configuration system that allows users to customize all aspects of vector generation and visualization. This section describes the configuration system and its features.

\subsection{VectorConfig Class}

The configuration system is implemented through the \texttt{VectorConfig} class, which provides a unified way to configure all aspects of vector generation and visualization. The class has the following attributes:

\begin{itemize}
    \item \texttt{origin}: The origin point for vector generation (default: [0, 0, 0]).
    \item \texttt{d}: The distance parameter for vector generation (default: 1.0).
    \item \texttt{theta}: The angle parameter for vector generation (default: $\pi/4$).
    \item \texttt{plot\_type}: The type of plot, either "3d" or "2d" (default: "3d").
    \item \texttt{title}: The title of the plot (default: None, which uses a default title based on the plot type).
    \item \texttt{show\_plot}: Whether to show the plot interactively (default: True).
    \item \texttt{save\_path}: The path to save the plot (default: None, which doesn't save the plot).
    \item \texttt{enhanced\_visualization}: Whether to use enhanced visualization features (default: True).
    \item \texttt{axis\_colors}: Custom colors for the X, Y, and Z axes (default: ['r', 'g', 'b']).
    \item \texttt{show\_coordinate\_labels}: Whether to show coordinate labels on the axes (default: True).
    \item \texttt{equal\_aspect\_ratio}: Whether to maintain an equal aspect ratio for 3D plots (default: True).
\end{itemize}

The class provides methods for initializing the configuration, saving it to a file, and loading it from a file.

\subsection{Initializing a Configuration}

A configuration can be initialized with default values or with custom values:

\begin{lstlisting}[language=Python]
# Initialize with default values
config = VectorConfig()

# Initialize with custom values
config = VectorConfig(
    origin=[1, 1, 1],
    d=2.0,
    theta=math.pi / 3,
    plot_type="2d",
    title="Vector Configuration",
    show_plot=False,
    save_path="custom.png",
    enhanced_visualization=True,
    axis_colors=['red', 'green', 'blue'],
    show_coordinate_labels=True,
    equal_aspect_ratio=True
)
\end{lstlisting}

\subsection{Using a Configuration}

A configuration can be used to generate and visualize orthogonal vectors:

\begin{lstlisting}[language=Python]
# Generate orthogonal vectors using the configuration
vectors = create_orthogonal_vectors(
    origin=config.origin,
    d=config.d,
    theta=config.theta
)

# Plot the vectors using the configuration
plot_vectors(
    vectors,
    origin=config.origin,
    plot_type=config.plot_type,
    title=config.title,
    show_plot=config.show_plot,
    save_path=config.save_path
)
\end{lstlisting}

\subsection{Saving a Configuration to a File}

A configuration can be saved to a JSON file for later use:

\begin{lstlisting}[language=Python]
# Save the configuration to a file
config.save_to_file("config.json")
\end{lstlisting}

The saved file will contain all the configuration parameters in JSON format:

\begin{lstlisting}[language=JSON]
{
    "origin": [1, 1, 1],
    "d": 2.0,
    "theta": 1.0471975511965976,
    "plot_type": "2d",
    "title": "Custom Configuration",
    "show_plot": false,
    "save_path": "custom.png"
}
\end{lstlisting}

\subsection{Loading a Configuration from a File}

A configuration can be loaded from a JSON file:

\begin{lstlisting}[language=Python]
# Load the configuration from a file
config = VectorConfig.load_from_file("config.json")
\end{lstlisting}

This creates a new \texttt{VectorConfig} object with the parameters specified in the file.

\subsection{Configuration in the Command-line Interface}

The command-line interface provides options for configuring vector generation and visualization:

\begin{itemize}
    \item \texttt{--origin}: Specifies the origin point as three space-separated values.
    \item \texttt{--d}: Specifies the distance parameter.
    \item \texttt{--theta}: Specifies the angle parameter in radians.
    \item \texttt{--plot-type}: Specifies the type of plot, either "3d" or "2d".
    \item \texttt{--title}: Specifies the title of the plot.
    \item \texttt{--no-show}: Prevents the plot from being displayed interactively.
    \item \texttt{--save-path}: Specifies the path to save the plot.
    \item \texttt{--config}: Specifies a configuration file to load.
    \item \texttt{--save-config}: Specifies a file to save the configuration to.
    \item \texttt{--no-enhanced-visualization}: Disables enhanced visualization features.
    \item \texttt{--axis-colors}: Specifies custom colors for the X, Y, and Z axes as three space-separated values.
    \item \texttt{--no-coordinate-labels}: Disables coordinate labels on the axes.
    \item \texttt{--no-equal-aspect-ratio}: Disables equal aspect ratio for 3D plots.
\end{itemize}

These options allow users to customize the configuration without modifying the code.

\subsection{Configuration File Format}

The configuration file is a JSON file with the following structure:

\begin{lstlisting}[language=JSON]
{
    "origin": [x, y, z],
    "d": float,
    "theta": float,
    "plot_type": "3d" or "2d",
    "title": string or null,
    "show_plot": boolean,
    "save_path": string or null,
    "enhanced_visualization": boolean,
    "axis_colors": ["r", "g", "b"],
    "show_coordinate_labels": boolean,
    "equal_aspect_ratio": boolean
}
\end{lstlisting}

All fields are optional and will use default values if not specified.

\subsection{Default Configuration}

The default configuration is as follows:

\begin{itemize}
    \item \texttt{origin}: [0, 0, 0]
    \item \texttt{d}: 1.0
    \item \texttt{theta}: $\pi/4$ (approximately 0.7853981633974483)
    \item \texttt{plot\_type}: "3d"
    \item \texttt{title}: None (uses a default title based on the plot type)
    \item \texttt{show\_plot}: True
    \item \texttt{save\_path}: None (doesn't save the plot)
    \item \texttt{enhanced\_visualization}: True
    \item \texttt{axis\_colors}: ['r', 'g', 'b']
    \item \texttt{show\_coordinate\_labels}: True
    \item \texttt{equal\_aspect\_ratio}: True
\end{itemize}

This configuration generates three orthogonal vectors from the origin [0, 0, 0] with a distance parameter of 1.0 and an angle parameter of $\pi/4$, and visualizes them in 3D.
