\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{physics}
\usepackage{tikz}




\usepackage{physics}

\usepackage{bm}
\usepackage{booktabs}
\usepackage{siunitx}

\usepackage{listings}
\usepackage{xcolor}

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\lstdefinestyle{mystyle}{
    backgroundcolor=\color{backcolour},   
    commentstyle=\color{codegreen},
    keywordstyle=\color{codepurple},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codegreen},
    basicstyle=\ttfamily\footnotesize,
    breakatwhitespace=false,         
    breaklines=true,                 
    captionpos=b,                    
    keepspaces=true,                 
    numbers=left,                    
    numbersep=5pt,                  
    showspaces=false,                
    showstringspaces=false,
    showtabs=false,                  
    tabsize=2
}

\lstset{style=mystyle}


\title{Notes -- WIP}
\date{\today}

\begin{document}

\maketitle

\section{Introduction}

Ez a dokumentum azt hivatott összegezni, amit a kollektív kónikus kereszteződési pontokkal kapcsolatos topologikus modellünkkel eddig elértünk.

Az alapelrendezése a rendszerünknek, hogy cavity-ben, három azonos kétatomos molekulát helyezünk el egy derékszögű koordináta rendszerben, ahol a molekulák tengelyei párhuzamosak az $x$, $y$ és $z$ tengelyekkel. Ez az elrendezés alapvetően megteremtheti a CI-k kialakulását a rendszerben.

\section{Model Potentials}

Our system is described by two potential energy surfaces:
\begin{align}
V_x(x) &= a_{V_x} \cdot x^2 \\
V_a(x) &= a_{V_a} \cdot (x - x_{\text{shift}})^2 + c_0
\end{align}

Where:
\begin{itemize}
    \item $a_{V_x}$ and $a_{V_a}$ are scaling parameters
    \item $x_{\text{shift}}$ is a shift parameter for the second potential
    \item $c_0$ is a constant energy offset
\end{itemize}

\section{Finding Conical Intersection Points}

A conical intersection occurs when the two potential energy surfaces meet at a point. For our model, this happens when:
\begin{equation}
V_a(x) - V_x(x) = 0
\end{equation}

\subsection{Mathematical Derivation}

Starting with the difference between the potentials:
\begin{align}
V_a(x) - V_x(x) &= a_{V_a} \cdot (x - x_{\text{shift}})^2 + c_0 - a_{V_x} \cdot x^2 \\
&= a_{V_a} \cdot x^2 - 2 \cdot a_{V_a} \cdot x_{\text{shift}} \cdot x + a_{V_a} \cdot x_{\text{shift}}^2 + c_0 - a_{V_x} \cdot x^2 \\
&= (a_{V_a} - a_{V_x}) \cdot x^2 - 2 \cdot a_{V_a} \cdot x_{\text{shift}} \cdot x + c_1
\end{align}

Where $c_1 = a_{V_a} \cdot x_{\text{shift}}^2 + c_0$.

We can rewrite this in the form of a perfect square plus a constant:
\begin{equation}
V_a(x) - V_x(x) = (a_{V_a} - a_{V_x}) \cdot (x - x_{\text{prime}})^2 + c_1'
\end{equation}

Where $x_{\text{prime}}$ is the point where the quadratic term has its minimum:
\begin{equation}
x_{\text{prime}} = \frac{a_{V_a}}{a_{V_a} - a_{V_x}} \cdot x_{\text{shift}}
\end{equation}

\section{Geometric Configuration of CI Points}

For a 3D system with coordinates $(r_0, r_1, r_2)$, we need to find points where the potentials are equal. This leads to a specific geometric arrangement.

\subsection{Key Insight}

If we have a point $R_0 = (r_0, r_1, r_2)$ where two of the coordinates give equal potential differences, then we can find a CI point. Specifically, if:
\begin{equation}
(V_a - V_x)(r_1) = (V_a - V_x)(r_2)
\end{equation}

Then we can express this in terms of distances from $x_{\text{prime}}$:
\begin{align}
r_1 &= x_{\text{prime}} - \delta \\
r_2 &= x_{\text{prime}} + \delta
\end{align}

Where $\delta$ is some displacement.

\subsection{Relationship Between Coordinates}

Given that only two coordinates can be equal at the same time to create a CI point, we can derive the following relationships:

If we set $r_0 = x_{\text{prime}} - \alpha$ for some value $\alpha$, then:
\begin{equation}
r_2 - r_0 = -2 \cdot (r_1 - r_0)
\end{equation}

This leads to:
\begin{align}
2(r_0 - r_1) &= r_2 - r_0 \\
2r_0 - 2x_{\text{prime}} + 2\delta &= x_{\text{prime}} + \delta - r_0 \\
3r_0 &= 3x_{\text{prime}} - 3\delta \\
\delta &= 3(x_{\text{prime}} - r_0)
\end{align}

\subsection{Expressions for $r_1$ and $r_2$}

Using the value of $\delta$, we can express $r_1$ and $r_2$ in terms of $r_0$ and $x_{\text{prime}}$:
\begin{align}
r_1 &= x_{\text{prime}} - \delta \\
&= x_{\text{prime}} - 3(x_{\text{prime}} - r_0) \\
&= 3r_0 - 2x_{\text{prime}} \\
&= r_0 - 2(x_{\text{prime}} - r_0)
\end{align}

Similarly:
\begin{align}
r_2 &= x_{\text{prime}} + \delta \\
&= x_{\text{prime}} + 3(x_{\text{prime}} - r_0) \\
&= 4x_{\text{prime}} - 3r_0 \\
&= r_0 + 4(x_{\text{prime}} - r_0)
\end{align}

\section{Distance to CI Points}

The distance from $R_0$ to the CI point can be calculated as:
\begin{align}
d_{\text{CI}} &= (r_2 - r_0) \cdot \frac{\sqrt{6}}{2} \\
&= 4(x_{\text{prime}} - r_0) \cdot \frac{\sqrt{6}}{2} \\
&= 2\sqrt{6} \cdot (x_{\text{prime}} - r_0)
\end{align}

\section{Implementation in Our Code}

In our code, we implement this by setting:
\begin{align}
r_0 &= x_{\text{prime}} \\
x &= 2(x_{\text{prime}} - r_0) = 0
\end{align}

For the CI point at index $n_{\text{CI}}$, we set:
\begin{equation}
R_0[i] = 
\begin{cases}
r_0 + x + x & \text{if } i = n_{\text{CI}} \\
r_0 - x & \text{otherwise}
\end{cases}
\end{equation}

With our parameter values, this gives $R_0 = [0.433, 0.433, 0.433]$ when $n_{\text{CI}} = 3$ (i.e., no specific CI point selected). We should run both a $\theta^*$- and a $d^*$-scan on this specific point.

\subsection{Arrowhead Hamiltonian}

Since the cavity excites collectively, the system is described by a 4×4 arrowhead Hamiltonian matrix with the following structure:

\begin{equation}
H(\bm{R}_{\theta}) = 
\begin{pmatrix}
\hbar\omega + \sum_i V_x(\bm{R}_i) & t_{01} & t_{02} & t_{03} \\
t_{01} & V_e(\bm{R}_0) & 0 & 0 \\
t_{02} & 0 & V_e(\bm{R}_1) & 0 \\
t_{03} & 0 & 0 & V_e(\bm{R}_2)
\end{pmatrix}
\end{equation}

where:
\begin{itemize}
    \item $\hbar\omega$ is the energy quantum of the system
    \item $V_x(\bm{R}_i) = a_{Vx} |\bm{R}_i|^2$ is the potential energy function
    \item $V_e(\bm{R}_i) = \sum_j V_x(\bm{R}_j) + V_a(\bm{R}_i) - V_x(\bm{R}_i)$ is the effective potential
    \item $V_a(\bm{R}_i) = a_{Va} (|\bm{R}_i - x_{\text{shift}}|^2 + c)$ is the additional potential
    \item $t_{0i}$ are the transition dipole moments between states
\end{itemize}

The parameter vector $\bm{R}_\theta$ traces a perfect circle orthogonal to the $x=y=z$ line in 3D space, parameterized by angle $\theta$.

\subsection{Berry Connection and Berry Phase}

The Berry connection $\bm{A}_n(\bm{R})$ for the $n$-th eigenstate is defined as:

\begin{equation}
\bm{A}_n(\bm{R}) = i\langle n(\bm{R})|\nabla_{\bm{R}}|n(\bm{R})\rangle
\end{equation}

The Berry phase $\gamma_n$ is the integral of the Berry connection around a closed loop $C$ in parameter space:

\begin{equation}
\gamma_n = \oint_C \bm{A}_n(\bm{R}) \cdot d\bm{R}
\end{equation}

\subsection{Off-Diagonal Elements}

The off-diagonal elements $\tau_{nm}(\theta)$ for $n \neq m$ provide information about the coupling between different eigenstates. These elements can be used to analyze transitions between states and to identify regions in parameter space where eigenstates are strongly coupled.


\section{The Tau Matrix Method}

We call this the "tau matrix method" because the key matrix is the tau matrix in the codebase. However, this matrix is actually the matrix of the Non-adiabatic coupling terms.

In this method, we first compute $\tau$ matrix, then use it to compute the Berry phase.

\subsection{Definition of the Tau Matrix}

This method is a numerical approach for computing the Berry phase. The key element is the tau matrix, whose elements are defined as:

\begin{equation}
\tau_{nm}(\theta) = \langle \psi_m(\theta) | \frac{\partial}{\partial\theta} | \psi_n(\theta) \rangle
\end{equation}

where $|\psi_n(\theta)\rangle$ is the $n$-th eigenstate of the Hamiltonian $H(\bm{R}_\theta)$.

\section{Numerical Results and Analysis}

\subsection{Berry Phase Matrix Structure}

The numerical results show that the Berry phase matrix $\gamma_{nm}$ has a specific structure with the following features:

\begin{itemize}
    \item The diagonal elements $\gamma_{nn}$ are typically small (order $10^{-3}$ to $10^{-4}$), indicating minimal geometric phase accumulation for individual states.
    
    \item Large off-diagonal elements (order $10^0$) appear in specific positions, particularly between states 1-2, 2-1, 1-3, and 3-1.
    
    \item The values close to $\pm 2\pi$ in the [1,2] and [2,1] positions indicate complete phase winding, which is a signature of non-trivial topology.
\end{itemize}

\subsection{Physical Interpretation}

The observed Berry phase structure has important physical implications:

\begin{itemize}
    \item The system exhibits non-trivial geometric phase accumulation primarily in the subspace spanned by states 1 and 2.
    
    \item The antisymmetric nature of the off-diagonal elements ($\gamma_{12} \approx -\gamma_{21}$) is consistent with the expected behavior of a quantum system with time-reversal symmetry.
    
    \item The $2\pi$ phase difference indicates that the system's parameter space trajectory encloses a topological singularity.
\end{itemize}


\end{document}
